---
title: "Figures projet Myco Phospho"
author: "Marie Locard-Paulet"
date: '`r date()`'
output: html_document
---

```{r packages, echo = F, message=FALSE}
require(ggplot2)
require(reshape2)
require(knitr)
require(gplots)


knitr::opts_chunk$set(echo = F)
source("RScripts/Functions/pdfMLP.R")
```

Data from the analysis described in the document `DataInspection_20171205`.

```{r input}
load("RData/Stat.RData")
tab$PhosID <- gsub("MSMEG_4757", "Fas1", as.character(tab$PhosID))
tab$PhosID <- gsub("FAS-", "Fas1-", as.character(tab$PhosID))
```

# Heatmap of most regulated proteins

I keep the sites that are significantly regulated in K Vs P and L Vs P.

```{r}
# Remove duplicate IDs for PknB from the 2 different species:
temp <- names(table(as.character(tab$PhosID)))[table(as.character(tab$PhosID)) > 1]
keepForPlot <- paste0(tab$PhosID[tab$PhosID %in% temp], "|", tab$ProtMatchingName[tab$PhosID %in% temp])
tab$PhosID[tab$PhosID %in% temp] <- paste0(tab$PhosID[tab$PhosID %in% temp], "|", tab$ProtMatchingName[tab$PhosID %in% temp])

# Select values:
int <- tab[tab$Significant_KVsP & tab$Significant_LVsP,grepl("MeanTech", names(tab))]
int$Phosphosite <- tab$PhosID[tab$Significant_KVsP & tab$Significant_LVsP]
names(int)[1:(ncol(int)-1)] <- substr(names(int), 11, 12)[1:(ncol(int)-1)]
int$RegulationDirection <- ifelse(tab$Intensity.K_Vs_Intensity.P_FC[tab$Significant_KVsP & tab$Significant_LVsP]>0, "Negative", "Positive")

# Order sites:
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%
int$Phosphosite <- factor(as.character(int$Phosphosite), levels = int$Phosphosite[order(tab$Intensity.K_Vs_Intensity.P_FC[tab$Significant_KVsP & tab$Significant_LVsP])]) # There are two entries with "mmpL5-T10"!! The sequences are different. Check what this is!!! Leading proteins are: I7G3R8 (Mmpl5 according to Uniprot) and I7GBH7 (also Mmpl5 according to Uniprot). These are identified in separate databases. I label them distinctly with their protein IDs
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%
int <- int[order(tab$Intensity.K_Vs_Intensity.P_FC[tab$Significant_KVsP & tab$Significant_LVsP]),]

# Select the top regulated sites:
perc <- 0.1 # Percentage of plotted sites
pos <- floor(length(int$RegulationDirection[int$RegulationDirection=="Positive"])*perc)
neg <- ceiling(length(int$RegulationDirection[int$RegulationDirection=="Negative"])*1)
intpos <- int[1:pos,1:10]
intneg <- int[(nrow(int)-neg+1):nrow(int),1:10]

# Create table for plotting:
gtabpos <- melt(intpos)
gtabneg <- melt(intneg)



colours = c("darkred","red", "gold")

gtabpos$Phosphosite <- factor(as.character(gtabpos$Phosphosite), levels = rev(levels(gtabpos$Phosphosite)))
g1 <- ggplot(gtabpos, aes(y = Phosphosite, x = variable)) + theme_minimal() + geom_tile(aes(fill = value)) + scale_fill_gradientn(colours = colours) + theme(axis.title.x = element_blank()) + theme(axis.ticks.x = element_blank(), axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 6)) + geom_vline(xintercept = c(3.5,6.5), size = 1) + ggtitle("Mean log2(LFQ intensities)")
print(g1)

g2 <- ggplot(gtabneg, aes(y = Phosphosite, x = variable)) + theme_minimal() + geom_tile(aes(fill = value)) + scale_fill_gradientn(colours = colours) + theme(axis.title.x = element_blank()) + theme(axis.ticks.x = element_blank(), axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 6)) + geom_vline(xintercept = c(3.5,6.5), size = 1) + ggtitle("Mean log2(LFQ intensities)")
print(g2)

pdfMLP(list(g1, g2), "Figures/HeatmapRegSites.pdf", "portrait")
```

# Proportion of phosphorylated amino-acids

```{r}
table(tab$Amino.acid)
table(tab$Amino.acid[tab$Significant_KVsL])
table(tab$Amino.acid[tab$Significant_KVsP])
table(tab$Amino.acid[tab$Significant_LVsP])

gtab <- data.frame("amino-acid" = c(c("S", "T", "Y"), "T", rep(c("S", "T", "Y"), 2)), "Count" = c(table(tab$Amino.acid), table(tab$Amino.acid[tab$Significant_KVsL]), table(tab$Amino.acid[tab$Significant_KVsP]), table(tab$Amino.acid[tab$Significant_LVsP])), "Conditions" = c(rep(c("Total regulated"), 3), "Regulated in KVsL", rep(c("Regulated in KVsP", "Regulated in LVsP"), each = 3)))

gtab$amino.acid <- factor(as.character(gtab$amino.acid), levels = c("Y", "S", "T"))

g <- ggplot(gtab, aes(x = Conditions, y = Count, fill = amino.acid)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("darkred", "chocolate1", "cornflowerblue")) + theme_light() + labs("title" = "Proportion of phosphorylated-serines and threonines") + xlab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(g)
pdfMLPManual(g, "Figures/AminoAcids.pdf", 5.5, 5)
```

# Global representation of the impact of PknB over-expression on the phosphoproteome


```{r}
MeanK <- rowMeans(tab[,289:291], na.rm = T)
MeanL <- rowMeans(tab[,292:294], na.rm = T)
MeanP <- rowMeans(tab[,295:297], na.rm = T)

gtab <- data.frame(MeanK, MeanL, MeanP)
g1 <- ggplot(gtab, aes(x = MeanK, y = MeanP)) + geom_abline(slope = 1, intercept = 0) + geom_point(alpha = 0.6) + ylab("Mean of log2(MS values) in PknB") + xlab("Mean of log2(MS values) in PknB-K40M") + theme_light()
g2 <- ggplot(gtab, aes(x = MeanL, y = MeanP)) + geom_abline(slope = 1, intercept = 0) + geom_point(alpha = 0.6) + xlab("Mean of log2(MS values) in the control") + ylab("Mean of log2(MS values) in PknB") + theme_light()
g3 <- ggplot(gtab, aes(x = MeanK, y = MeanL)) + geom_abline(slope = 1, intercept = 0) + geom_point(alpha = 0.6) + xlab("Mean of log2(MS values) in PknB-K40M") + ylab("Mean of log2(MS values) in the control") + theme_light()

plot(g1); plot(g2); plot(g3)
pdfMLPManual(list(g1, g2, g3), "Figures/CorrelationsPhospho.pdf", 8,7)

gtab <- data.frame("MeanK" = table(is.na(MeanK)), "MeanP" = table(is.na(MeanP)), "MeanL" = table(is.na(MeanL)))
gtab <- melt(gtab)
gtab$Detected <- ifelse(gtab$MeanL.Var1 == "TRUE", "No", "yes")
gtab$variable <- gsub(".Freq", "", as.character(gtab$variable), fixed = T)
g <- ggplot(gtab[gtab$Detected=="yes",], aes(y= value, x = variable)) + geom_bar(stat = "identity", position = "dodge", alpha = 0.8, col = "black") + theme_light() + ylab("Count") + xlab("") + ggtitle("Number of phosphorylation site")
print(g)

pdfMLPManual(g, "Figures/NumberIDPhospho.pdf", 7,6)
```

Venn diagrams of detected sites:

```{r}
ID_K <- tab$PhosID[!is.na(MeanK)]
ID_P <- tab$PhosID[!is.na(MeanP)]
ID_L <- tab$PhosID[!is.na(MeanL)]
venn(list("K"=ID_K, "P" = ID_P, "L" = ID_L))
title("Identified phosphosites")
```

and of regulated sites:

```{r}
KP <- tab$PhosID[tab$Significant_KVsP]
KL <- tab$PhosID[tab$Significant_KVsL]
LP <- tab$PhosID[tab$Significant_LVsP]
venn(list("KVsP"=KP, "KVsL" = KL, "LVsP" = LP))
title("Regulated phosphosites")
```


# Heatmap with the proteins of interest:

```{r}
IntProt <- read.table("OutputTables/ProteinsOfInterest3.txt", sep = "\t", header = T)
keep <- ((toupper(tab$GeneID2) %in% toupper(c(as.character(IntProt$Entry.name), "pknb", "pknbtub", "mtpknb", keepForPlot))) | tab$GeneID2 == "MSMEG_4757")


# Select values:
int <- tab[keep,grepl("Intensity.K_Vs_Intensity.P_FC", names(tab))]
int <- data.frame("FC_KVsP"= int, "Phosphosite" = tab$PhosID[keep], stringsAsFactors = F)

int <- int[!is.na(int$Phosphosite),]

# Order sites:
int$Phosphosite <- factor(as.character(int$Phosphosite), levels = sort(int$Phosphosite))
int <- int[order(int$Phosphosite),]

colours = c("darkred","red", "gold")
int$x <- rep(1, nrow(int))


g1 <- ggplot(int, aes(y = Phosphosite, x = x)) + 
  theme_minimal() + 
  geom_tile(aes(fill = -FC_KVsP)) + 
  scale_fill_gradientn(colours = colours) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.ticks.x = element_blank(), axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 6)) + 
  labs(title = "Log2-transformed Fold changes\nof sites of interest", subtitle = "P Vs K")
print(g1)
```


# Heatmap with Fold changes for the pathway figure in the paper:

```{r}
IntProt <- read.table("OutputTables/ProteinsOfInterest3.txt", sep = "\t", header = T)
IntProt <- IntProt[IntProt$keep,]
tab$GeneID2 <- gsub(" ", "", tab$GeneID2)
tab$PhosID <- gsub(" ", "", tab$PhosID)

keep <- ((toupper(tab$GeneID2) %in% toupper(c(as.character(IntProt$Entry.name), "pknb", "pknbtub", "mtpknb", keepForPlot))) | tab$GeneID2 == "MSMEG_4757")


# Distinction between the tub Vs smeg PknB:
tab$GeneID2[tab$ProtMatchingName=="A0QNG1" & !is.na(tab$ProtMatchingName)] <- paste0("Smeg_", tab$GeneID2[tab$ProtMatchingName=="A0QNG1" & !is.na(tab$ProtMatchingName)])
tab$PhosID[tab$ProtMatchingName=="A0QNG1" & !is.na(tab$ProtMatchingName)] <- paste0("Smeg_", tab$PhosID[tab$ProtMatchingName=="A0QNG1" & !is.na(tab$ProtMatchingName)])

# Select values:
int <- tab[keep,grepl("MeanTech", names(tab))]
int$Phosphosite <- tab$PhosID[keep]

names(int)[1:(ncol(int)-1)] <- substr(names(int), 11, 12)[1:(ncol(int)-1)]

int$RegulationDirection <- ifelse(tab$Intensity.K_Vs_Intensity.P_FC[keep]>0, "Negative", "Positive")
int <- int[!is.na(int$Phosphosite),]

# Order sites:
int$Phosphosite <- factor(as.character(int$Phosphosite), levels = int$Phosphosite[order(tab$Intensity.K_Vs_Intensity.P_FC[keep])])
int <- int[order(tab$Intensity.K_Vs_Intensity.P_FC[keep]),]

#intpos <- int[int$RegulationDirection=="Positive",]
intpos <- int
intneg <- int[int$RegulationDirection=="Negative",]

intpos <- intpos[!is.na(intpos$Phosphosite),]
intneg <- intneg[!is.na(intneg$Phosphosite),]


# Create table for plotting:
gtabpos <- melt(intpos)
gtabneg <- melt(intneg)

# For marginal plots:
library(scales)
require(gtable)

## To format heatmap y.ticks with appropriate width (5 chars),
## to align with gg_rows y.tics
ytickform <- function(x){
    lab <- sprintf("%05s",x)
}

colours = c("darkred","red", "gold")

#gtabpos <- gtabpos[!is.na(gtabpos$value),]
gtabpos$Phosphosite <- factor(as.character(gtabpos$Phosphosite), levels = rev(levels(gtabpos$Phosphosite)))
gtabpos$variable <- factor(as.character(gtabpos$variable), levels = c("L1", "L2", "L3", "K1", "K2", "K3", "P1", "P2", "P3"))

# Table with FCs:
dfFC <- data.frame("FC" = -tab$Intensity.K_Vs_Intensity.P_FC, "Phosphosite" = tab$PhosID, "Significative" = tab$Significant_KVsP)
#dfFC <- dfFC[order(dfFC$FC, decreasing = T),]
#dfFC <- dfFC[!duplicated(dfFC$Phosphosite),]
dfFC <- dfFC[dfFC$Phosphosite %in% as.character(gtabpos$Phosphosite),]
dfFC <- dfFC[!is.na(dfFC$Phosphosite),]
dfFC <- dfFC[match(as.character(gtabpos$Phosphosite), dfFC$Phosphosite),]
dfFC <- dfFC[!duplicated(dfFC),]
dfFC$Phosphosite <- factor(as.character(dfFC$Phosphosite), levels = as.character(dfFC$Phosphosite[order(dfFC$FC, decreasing = F)]))
dfFC$Gene <- sapply(as.character(dfFC$Phosphosite), function(x) {
  strsplit(x, "-", fixed = T)[[1]][1]
})
dfFC$Gene <- gsub("Smeg_", "", dfFC$Gene) 
dfFC$Category <- as.character(IntProt$category[match(toupper(dfFC$Gene), toupper(IntProt$Entry.name))])
dfFC$Category[dfFC$Gene=="Fas1"] <- IntProt$category[IntProt$Entry.name=="FAS"]
dfFC$Category[is.na(dfFC$Category)] <- "PknB"
dfFC$Category[dfFC$Category=="1"] <- "Fatty acid synthesis, elongation and modification"

dfFC <- dfFC[!is.na(dfFC$FC),]

# I loop over the categories:

cat <- as.character(unique(dfFC$Category))
require(grid)

# Homogenizing colour scale:
val1 <- min(gtabpos$value, na.rm = T)
val4 <- max(gtabpos$value, na.rm = T)
val <-  (max(gtabpos$value, na.rm = T) - min(gtabpos$value, na.rm = T))/3
val2 <- val1+val
val3 <- val2+val
mid <- (max(gtabpos$value, na.rm = T) - min(gtabpos$value, na.rm = T))/2+val1

for (el in cat) {
  gtab <- gtabpos[gtabpos$Phosphosite %in% dfFC$Phosphosite[dfFC$Category==el],]

  
  #gtab <- gtab[!duplicated(gtab$Phosphosite),]

  g1 <- ggplot(gtab, aes(y = Phosphosite, x = variable)) + theme_minimal() + geom_tile(aes(fill = value)) + scale_fill_gradientn(colours = colours, limits = c(val1, val4)) + theme(axis.title.x = element_blank()) + theme(axis.ticks.x = element_blank(), axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 6)) + geom_vline(xintercept = c(3.5,6.5), size = 1) + ggtitle(el)
  print(g1)
  
  gg_hm <- g1 + scale_y_discrete(label=ytickform) + theme(legend.position = "bottom", plot.margin = unit(c(3,3,3,3), "mm")) + theme(plot.margin = unit(c(3,3,3,3), "mm"))
  
  gg_rows <- ggplot(dfFC[dfFC$Phosphosite %in% dfFC$Phosphosite[dfFC$Category==el],], aes(y = FC, x = Phosphosite, fill = Significative)) + geom_bar(stat = "identity", col = "grey30") + scale_fill_manual(values = c("grey30", "steelblue1")) + coord_flip() + theme_minimal() + xlab("") + theme(axis.title.y=element_blank(), axis.text.y=element_blank(),axis.ticks.y=element_blank(), axis.title.x=element_blank(), plot.margin = unit(c(9,1,3,1), "mm"))
  
  ## extract legend from heatmap
  g <- ggplotGrob(gg_hm)$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  
  ## plot heatmap without legend
  g <- ggplotGrob(gg_hm + theme(legend.position="none"))
  
  ## add column and put column barplot within
  g <- gtable_add_cols(g, unit(5,"cm"))
  g <- gtable_add_grob(g, ggplotGrob(gg_rows), t = 1, l=ncol(g), b=nrow(g), r=ncol(g))
  
  ## add row and put legend within
  g <- gtable_add_rows(g, unit(1.1,"cm"))
  g <- gtable_add_grob(g, legend, t = nrow(g), l=1, b=nrow(g), r=ncol(g)-1)
  
  name <- gsub(",", "", el)
  name <- paste0("Figures/HeatmapRegSitesOfInterest_", name, ".pdf")
  pdf(name, useDingbats=FALSE, 8.27, 11.69)
  grid.newpage()
  grid.draw(g)
  dev.off()
}
```

********************************************************************************

```{r}
Sys.info()
```